The relationship between HbA1c, GMI, and average glucose in adults with CFRD

### 2025-08-12 Christine Chan/EnVision CF/cgm_and_a1c.qmd
formatting for HbA1c and CGM

```{r setup}
#| include: false
library(tidyverse)
library(redcapAPI)
library(gtsummary)
library(labelled)
library(ggpmisc)
library(blandr)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| include: false
# Import from REDCap
unlockREDCap(c(rcon = "EnVision CF HbA1c and CGM"),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ahc.umn.edu/redcap/api/"
)
df <- exportRecordsTyped(rcon)
# Start building analysis dataset
df <- df %>%
  # Format columns
  mutate(
    # Numeric
    across(c(age, bmi, cfrd_diagnosis_age, average_sensor_glucose), ~
      as.numeric(.)),
    across(c(hb_a1c_value, gmi), ~
      as.numeric(sub("%", "", .))),
    # Categorical
    combined_race = factor(race,
      levels = c(
        "White", "Black or African American", "Asian", "More Than One Race",
        "Unknown / Not Reported"
      ),
      labels = c(
        "White", "Black or African American", "Other/Unknown", "Other/Unknown",
        "Other/Unknown"
      )
    )
  ) %>%
  # Drop unnecessary columns
  select(
    subject_id, age, cfrd_diagnosis_age, bmi, combined_race, sex, pi,
    cftr_modulator_use, type_of_modulator, type_of_cgm, hb_a1c_value, gmi,
    average_sensor_glucose
  ) %>%
  # Labels
  set_variable_labels(
    age = "Age (years)", cfrd_diagnosis_age = "Age at CFRD Diagnosis (years)",
    bmi = "BMI (kg/m2)", combined_race = "Race", sex = "Sex at Birth",
    type_of_modulator = "Type of Modulator", hb_a1c_value = "HbA1c (%)",
    gmi = "GMI (%)", average_sensor_glucose = "Mean Sensor Glucose (mg/dL)",
    cftr_modulator_use = "CFTR Modulator", pi = "Pancreatic Insufficiency"
  )
```

@@

```{r}
#| warning: false
ggplot(df, aes(x = hb_a1c_value, y = gmi)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2", "p", "n")))

blandr.draw(df$hb_a1c_value, df$gmi, ciDisplay = F, annotate = T)
```

```{r}
#| warning: false
ggplot(df, aes(x = hb_a1c_value, y = average_sensor_glucose)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2", "p", "n")))

blandr.draw(df$hb_a1c_value, df$average_sensor_glucose,
  ciDisplay = F, annotate = T
)
```

>>>>

```{r setup}
#| include: false
library(keyring)
library(tidyverse)
library(redcapAPI)
library(gtsummary)
library(labelled)
library(ggpmisc)
library(blandr)
home_dir <- switch(
    Sys.info()["sysname"],
    "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
    "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
    "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
github_dir <- switch(
    Sys.info()["sysname"],
    "Darwin" = "/Users/tim/Documents/GitHub",
    "Windows" = "C:/Users/Tim/Documents/GitHub",
    "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| include: false
# Import from REDCap
token <- key_get("envision_cf_hba1c_cgm")
rcon <- redcapConnection(
    url = "https://redcap.ahc.umn.edu/redcap/api/",
    token = token
)
df <- exportRecordsTyped(rcon)
# Start building analysis dataset
df <- df %>%
    # Format columns
    mutate(
        # Numeric
        across(
            c(age, bmi, cfrd_diagnosis_age, average_sensor_glucose),
            ~ as.numeric(.)
        ),
        across(c(hb_a1c_value, gmi), ~ as.numeric(sub("%", "", .))),
        # Categorical
        combined_race = factor(
            race,
            levels = c(
                "White",
                "Black or African American",
                "Asian",
                "More Than One Race",
                "Unknown / Not Reported"
            ),
            labels = c(
                "White",
                "Black or African American",
                "Other/Unknown",
                "Other/Unknown",
                "Other/Unknown"
            )
        )
    ) %>%
    # Drop unnecessary columns
    select(
        subject_id,
        age,
        cfrd_diagnosis_age,
        bmi,
        combined_race,
        sex,
        pi,
        cftr_modulator_use,
        type_of_modulator,
        type_of_cgm,
        hb_a1c_value,
        gmi,
        average_sensor_glucose
    ) %>%
    # Labels
    set_variable_labels(
        age = "Age (years)",
        cfrd_diagnosis_age = "Age at CFRD Diagnosis (years)",
        bmi = "BMI (kg/m2)",
        combined_race = "Race",
        sex = "Sex at Birth",
        type_of_modulator = "Type of Modulator",
        hb_a1c_value = "HbA1c (%)",
        gmi = "GMI (%)",
        average_sensor_glucose = "Mean Sensor Glucose (mg/dL)",
        cftr_modulator_use = "CFTR Modulator",
        pi = "Pancreatic Insufficiency"
    )
```

@@

```{r}
#| warning: false
ggplot(df, aes(x = gmi, y = hb_a1c_value)) +
    geom_point() +
    geom_smooth(method = "lm", se = F) +
    stat_poly_line() +
    stat_poly_eq(use_label(c("eq", "R2", "p", "n")))

blandr.draw(df$gmi, df$hb_a1c_value, ciDisplay = F, annotate = T)
```

# Mean sensor glucose and HbA1c

```{r}
#| warning: false
ggplot(df, aes(x = average_sensor_glucose, y = hb_a1c_value)) +
    geom_point() +
    geom_smooth(method = "lm", se = F) +
    stat_poly_line() +
    stat_poly_eq(use_label(c("eq", "R2", "p", "n")))
```
